\begin{titlepage}

\title{Sweet and Timely Insurance: \\
The Role of Honey in Reducing Coffee Producer Food
Insecurity Exposure in Mexico}
\author{Grant X. Storer\thanks{Graduate student, Department of Economics, University of San Francisco. email: \protect\href{gxstorer@usfca.edu}. Address: 2130 Fulton St., San Francisco, CA 94117. Phone: (+1) 415-422-2765.}, Stephen Pitts\thanks{Ph.D. Candidate, Department of Applied Economics, University of Minnesota. Corresponding author. email: \protect\href{pitts071@umn.edu)}. Address: 1994 Buford Ave, St. Paul, MN 55108. Phone: (+1) 612-625-1222.}, Chris M. Boyd\thanks{Assistant Professor, Department of Economics, Towson University. email: \protect\href{cboydleon@towson.edu}. Address: 8000 York Road, Stephens Hall 101-N, Towson, MD 21252. Phone: (+1) 410-704-2959.} \ and\\ Jesse Anttila-Hughes\thanks{Associate Professor, Department of Economics, University of San Francisco. email:\protect\href{jkanttilahughes@usfca.edu}. Address: 2130 Fulton St., San Francisco, CA 94117. Phone: (+1) 415-422-2765.}}

\maketitle

\begin{abstract}
Smallholder coffee producers face a combination of pre- and post-harvest risk factors
that leaves them particularly vulnerable to food insecurity. A popular form of on-farm
diversification is honey production through beekeeping, that has both nutritional and
commercial value. This study investigates the role of honey production as means of food
security management due to the heightened pollinating activity during the coffee flowering
stage that follows the annual coffee harvest provides an additional non-contemporaneous
source of income. Using primary data collected in coffee-producing regions of Chiapas,
Mexico, I find that during the honey harvest months, which occurs during the early stage of
the lean season, beekeeping coffee producers are less exposed to food insecurity over coffee
producers who don’t diversify into honey. 

\end{abstract}

\vspace{1cm} 
\textbf{\small{}JEL Codes:}{\small{} D1, D6, D8, I3, O1, Q1.}{\small\par}

\textbf{\small{}Keywords:}{\small{} Coffee, Food Insecurity, Diversification, Honey, Beekeeping, Technology Adoption, Indigenous Communities, Chiapas, Mexico.}{\small\par}

\thispagestyle{empty} 

\end{titlepage}