---
title: "Probelem set-1 in R"
output:
  word_document: default
  html_notebook: default
---

```{r}
library(haven)
setwd("~/Dropbox/School/USF/Econometrics Causal/Original Data")
data <- read_dta("ps1_standard_errors.dta")
```

2. (Really) Understanding Robust and Clustered Standard Errors: 

a)	Using the data we practiced with in lecture, where consumption = (70, 65, 90, 95, 110, 115, 80, 200, 190, 100) 
and income = (80, 100, 120, 140, 160, 180, 200, 220, 240, 260), first simply code regress consumption income in 
Stata to get an OLS estimate of the relationship between consumption and income. 
```{r}

library("dplyr")
#Using rename()
data <- data %>% rename("consumtion" = "ys")
data <- data %>% rename("income" = "Xs")

```

simple regression 
```{r}
reg = lm(consumtion~income, data = data)
summary(reg)
```
b.  estimate for β ̂ = (X'X)^-1(X'y). (Don't forget to include the column of 1's so that you estimate an intercept.)
```{r}
data$residuals <- residuals(reg)
data$r_sqr ='^'(data$residuals,2)
data
```



```{r}
X <- cbind(data$ones, data$income)
y <- cbind(data$consumtion)
X
y
```
matrix Betas = (invsym(X' * X)) * (X' * y)
```{r}
betas <- (solve(t(X)%*%X))%*%((t(X))%*%y)
betas
```


c)manually calculate the OLS standard error for  β ̂.  Remember that 
var(β ̂) = s2(X'X)-1 , where s2= Sum(ei2/(n – k)). (Spoiler Alert: You should get 0.20827.) 

*First create the Buns to the Sandwich estimator  (X'X)-1. 2x2:
matrix Bun = invsym(X' * X)
matrix list Bun
```{r}
Bun <- solve(t(X)%*%X)
Bun
```
then the Meat… ei^2... 10x1
```{r}
E_squareds<- cbind(data$r_sqr)
```
s^2 
```{r}
s_sqr = sum(E_squareds)/(nrow(X)-ncol(X))
s_sqr
```
get variance β ̂= s2(X'X)^-1 . 2x2 
```{r}
var_B = s_sqr*Bun
var_B
```
getting standard errors  SE= (var(B^))^1/2
```{r}
SE_Bo=sqrt(var_B[1,1])
SE_B1=sqrt(var_B[2,2])
SE_Bo
SE_B1
```



